---
title: "Reproducible Research"
author: "Abinashi Prakash"
date: "27 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reproducible Research: Peer Assessment 1

##Loading and preprocessing the data
```{r ,echo=TRUE}

###1.Load the data 

source_data = read.csv("C:/Users/abinashi.prakash/Desktop/Coursera/repdata_data_activity/activity.csv")

###2.Process/transform the data (if necessary) into a format suitable for your analysis--Not required as data is already in required format.

```

##What is mean total number of steps taken per day?
```{r Steps per Day,echo=TRUE}

###1.Calculate the total number of steps taken per day

Tot_steps <- sum(source_data$steps, na.rm = TRUE)

Tot_steps_day <- aggregate(steps~date, data=source_data, FUN=sum, na.rm=TRUE)
```

```{r Histogram Steps Each Day,echo=TRUE}
###2. Make a histogram of the total number of steps taken each day

hist(Tot_steps_day$steps,main = "Total Steps per Day",xlab = "Number of Steps")

```

```{r,echo=TRUE}

###3.Calculate and report the mean and median of the total number of steps taken per day

Tot_steps_day_mean <- mean(Tot_steps_day$steps,na.rm = TRUE)
Tot_steps_day_mean

Tot_steps_day_median <- median(Tot_steps_day$steps,na.rm = TRUE)
Tot_steps_day_median

```


##What is the average daily activity pattern?

```{r Time Series Plot for 5 mins interval,echo=TRUE}

###1.Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

avg_interval <- aggregate(steps~interval, data=source_data, FUN=mean, na.rm=TRUE)

plot(x = avg_interval$interval, y = avg_interval$steps, type = "l") 

```

###2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r,echo=TRUE}
max_steps <- max(avg_interval$steps) ##Storing Max steps in a variable
for (i in 1:288) 
{
  if (avg_interval$steps[i] == max_steps)
    max_steps_five_min_interval <- avg_interval$interval[i]
}
max_steps_five_min_interval 

```


##Imputing missing values
```{r,echo=TRUE}

###1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
total_na <- 0
for (i in 1:17568)
{
  if(is.na(source_data$steps[i])) 
    total_na <- total_na+1 
}
total_na
```

```{r,echo=TRUE}
###2.Devise a strategy for filling in all of the missing values in the dataset.
###3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
source_data_filled_in <- source_data
for (i in 1:17568) 
{
  if(is.na(source_data_filled_in$steps[i])) # if steps is na store the pointer 
  { 
    five_min_point <- source_data_filled_in$interval[i] 
    for (j in 1:288)  
    {
      if (avg_interval$interval[j] == five_min_point) 
        source_data_filled_in$steps[i] <- avg_interval$steps[j] 
      
    }
  }
}
```


###4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r Histogram of number of steps taken each day without NA,echo=TRUE}

Tot_steps_day_filled_in <- aggregate(steps~date, data=source_data_filled_in, FUN=sum, na.rm=TRUE)
###Histogram of the total number of steps taken each day

hist(Tot_steps_day_filled_in$steps)

###Calculate and report the mean and median total number of steps taken per day. 
Tot_steps_day_mean_filled_in <- mean(Tot_steps_day_filled_in$steps)
Tot_steps_day_mean_filled_in

Tot_steps_day_median_filled_in <- median(Tot_steps_day_filled_in$steps)
Tot_steps_day_median_filled_in
```


##Are there differences in activity patterns between weekdays and weekends?

```{r,echo=TRUE}

###1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

week <- weekdays(as.Date(source_data_filled_in$date))
week_day <- week

for (i in 1:17568) 
{
  if(week[i] == "Sunday")
    week_day[i] <- 'weekend'
  if(week[i] == "Monday")
    week_day[i] <- 'weekday'
  if(week[i] == "Tuesday")
    week_day[i] <- 'weekday'
  if(week[i] == "Wednesday")
    week_day[i] <- 'weekday'
  if(week[i] == "Thursday")
    week_day[i] <- 'weekday'
  if(week[i] == "Friday")
    week_day[i] <- 'weekday'
  if(week[i] == "Saturday")
    week_day[i] <- 'weekend'
}

source_data_filled_in$weekday <- week_day
```

```{r,echo=TRUE}
###2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

weekday <- grep("weekday",source_data_filled_in$weekday)
wday_frame <- source_data_filled_in[weekday,]
wend_frame <- source_data_filled_in[-weekday,]

## Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, 
##    averaged across all days (yaxis)
avg_interval_wday <- aggregate(steps~interval, data=wday_frame, FUN=mean, na.rm=TRUE)
##avg_interval_wday
avg_interval_wend <- aggregate(steps~interval, data=wend_frame, FUN=mean, na.rm=TRUE)
##avg_interval_wend
```

```{r Plots for Weekdays,echo=TRUE}
### Plots for Weekend and Weekdays
plot(x = avg_interval_wday$interval, y = avg_interval_wday$steps, type = "l") 
```

```{r Plots for Weekend,echo=TRUE}
plot(x = avg_interval_wend$interval, y = avg_interval_wend$steps, type = "l") 
```

